A Transactional Memory System for Parallel Python

### 2014-06-07 talk/dls2014/paper/paper.tex
more updates

\newcommand{\mynote}[2]{%
  \textcolor{red}{%
    \fbox{\bfseries\sffamily\scriptsize#1}%
    {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}%
  }%
}

\newcommand\remi[1]{\mynote{Remi}{#1}}
\newcommand\cfbolz[1]{\mynote{cfbolz}{#1}}
\newcommand\arigo[1]{\mynote{arigo}{#1}}

\newcommand{\comment}[1]{}

@@

\comment{
A Platform for Parallelism in Dynamic languages
Parallelism for Python
A TM Implementation for Python
Transactional Memory as a foundation for parallel Python
Transactional Memory as a foundation for parallelism in Python
Parallel Python based on VM-assisted TM
Parallel Python based on VM-assisted Transactional Memory
A Platform for parallel Python
A TM Platform for parallel Python
A Transactional Memory Platform for parallel Python
A Transactional Memory system for parallel Python

Virtual Memory Assisted Transactional Memory for Dynamic Languages}

@@

This paper makes the following contributions:
\begin{itemize}[noitemsep]
\item We introduce a new STM system that performs well for dynamic
  language interpreters.
\item We integrate the system closely with a garbage collector
  (GC) to lower the overhead of STM.
\item This new STM system is used to replace the GIL in one
  implementation of Python and is then evaluated.
\end{itemize}
The STM described here is efficient and has low overhead and is
therefore beneficial also for environments with a small number of
threads or cores.

@@

\subsection{Architecture}

Our TM system is designed as a library that covers all aspects around
transactions and object management. It is designed for object-oriented
dynamic language VMs as a replacement for the GIL.

The library consists of two parts: (I) It provides a simple interface
to starting and committing transactions, as well as the required read
and write barriers. (II) It also includes a \emph{garbage collector
(GC)} that is closely integrated with the TM part.

@@

\begin{lstlisting}
void stm_start_transaction()
void stm_commit_transaction()
void stm_become_inevitable()
void stm_read(object_t *SO)
void stm_write(object_t *SO)
object_t *stm_allocate(size_t size_rounded)
STM_PUSH_ROOT(object_t *SO)
STM_POP_ROOT(object_t *SO)
\end{lstlisting}

@@

The next step involves using \lstinline!remap_file_pages()!, a Linux
system call\footnote{remap\_file\_pages() can also be done with more mmap()
calls in a roughly portable way, but the former is more efficient on Linux.},
to establish the \emph{fully-shared configuration}.  All pages
of segments $>0$ map to the pages of the sharing-segment.

However, the layout of a segment is not uniform and we actually
privatise a few areas again right away. These areas are illustrated in
Figure~\ref{fig:Segment-Layout} and explained here:
\begin{description}[noitemsep]
\item [{NULL~page:}] This page is unmapped and will produce a
  segmentation violation when accessed. We use this to detect
  erroneous dereferencing of \lstinline!NULL! references.  All
  $\%gs{::}SO$ translated to linear addresses will point to NULL pages
  if $SO$ is set to \lstinline!NULL!.
\item [{Segment-local~data:}] Some area private to the segment that
  contains segment-local information for bookkeeping.
\item [{Read~markers:}] These are private pages that store information about
  which objects were read in the current transaction running in this
  segment.
\item [{Nursery:}] This private area contains all the freshly allocated
  objects (\emph{young objects}) of the current transaction. The GC
  uses bump-pointer allocation in this area to allocate objects in the
  first generation.
\item [{Old~object~space:}] These pages are the ones that are really
  shared between segments. They mostly contain old objects but also
  some young ones that were too big to be allocated in the nursery.
\end{description}

@@

At the start of a transaction, the thread needs to acquire a
segment. It may have to wait until another thread finishes its
transaction and releases its segment. Fairness is not guaranteed yet, as
we simply assume a fair scheduling policy in the operating system when
waiting on a condition variable.

@@

To improve this situation even more, we introduce the concept of
\emph{overflow objects}. If a minor collection needs to occur during a
transaction, we empty the nursery and mark each surviving object in
the old object space with an \lstinline!overflow_number!  globally
unique to the current transaction. That way we can still detect in a
medium-fast path inside barriers that the object still belongs to the
current transaction. Other and later transactions will treat them
like regular \emph{old} objects.

@@

This area can be seen as a continuous, segment-local array of bytes
that is indexed with an object's reference ($SO$) divided by 16. So
that each object has its own read marker, all objects have a size of
at least 16 bytes. Otherwise there could be false conflicts when
reading from two adjacent objects that share a single marker.
Instead of just setting the byte to \lstinline!true!  if the
corresponding object was read, we set it to a \lstinline!read_version!
belonging to the transaction, which will be incremented on each
commit.  Thereby, we can avoid resetting the bytes to
\lstinline!false!  on commit and only need to do this every 255
transactions. The whole code for the barrier is easily optimisable for
compilers as well as perfectly predictable for
CPUs\footnote{Additional benefit: The read barrier is not constrained
to execute before the actual read -- both the compiler and the CPU are
free to reorder or group them.}:

\begin{lstlisting}
void stm_read(SO):
    *(SO >> 4) = read_version
\end{lstlisting}

@@

The job of the write barrier is twofold: first, it serves as a write
barrier for the garbage collector and second, it supports COW and adds
objects to the write set of the transaction. These two jobs are
depicted in Figure~\ref{fig:obj_states}.

The \textbf{fast path} of the write barrier is very simple. We only
need to check for the flag \lstinline!WRITE_BARRIER! in the object's
header through its $SO$ and call the slow path if it is set. This flag
is set either if the object is old and comes from an earlier
transaction, or if there was a minor collection which will add the
flag again on all objects and may produce overflow objects. The flag
is never set on freshly allocated objects, because they do not need
to be traced by the GC or handled by the STM system. This fast path
covers all cases for the GC and for the STM system.

\begin{lstlisting}
void stm_write(SO):
	if SO->flags & WRITE_BARRIER:
		write_slowpath(SO)
\end{lstlisting}

@@

The \textbf{slow path} is shown in Listing~\ref{lst:w_slow}.
First comes the \emph{GC part}: In any case, the object will be added
to the list of objects that need tracing in the next minor collection
(\lstinline!to_trace!).  This is necessary in case we write a
reference to it that points to a young object. We then need to trace
it during the next minor collection to mark the young object
alive and to update its reference to the new location it gets moved
to.

@@

For the \emph{STM part}, we first perform a read barrier on the
object. We then try to acquire its write lock. \lstinline!write_locks!
is a simple, \textbf{global} array of bytes that is indexed with the
$SO$ of the object divided by 16. Note, ``global'' here really means
it is a single array with data for all segments, there is no address
translation going on to access its elements contrary to e.g.\ the
read markers array.  If we already own the lock, we are done.
If someone else owns the lock, we perform  write-write contention
management that aborts either us or the current owner of the
object.  If we succeed in acquiring the lock using an atomic
\lstinline!cmp_and_swap!, we need to add the object to the write set
(a simple list called \lstinline!modified_old_objects!)  and privatise
all pages belonging to it (copy-on-write).

@@

Note that we have three kinds of objects: \emph{young, overflow} and
\emph{old} objects. Young and overflow objects were created in the
same transaction; old objects always come from previous transactions.
The close integration between STM and the GC allows us to use this
information effectively: young objects do not trigger the slow path at
all, overflow objects only need to be handled by the GC part of the
write barrier once between collections, and only the old objects need
to be handled by the STM part \textbf{once per transaction}. This keeps the
overhead of the write barrier very low.

@@

A requirement for performing a commit is to synchronise all threads so
that we can safely update objects in other segments. To make this
synchronisation fast and cheap, we do not want to insert an additional
check regularly to see if synchronisation is requested. We
use a trick relying on the fact that dynamic languages are usually
very high-level and thus allocate a lot of objects very regularly.
This is done through the function \lstinline!stm_allocate!  shown
below:

\begin{lstlisting}
object_t *stm_allocate(ssize_t size_rounded):
    result = nursery_current
	nursery_current += size_rounded
	if nursery_current > nursery_end:
		return allocate_slowpath(size_rounded)
	return result
\end{lstlisting}

@@

On encountering conflicts, we employ contention management to solve
the problem as well as we can. The general rules are:
\begin{itemize}[noitemsep]
\item prefer transactions that started earlier to younger transactions
  to let long transactions succeed more often
\item to support \emph{inevitable} transactions, we always prefer them
  to others since they cannot abort (similar to~\cite{blundell06})
\end{itemize}
We can either simply abort a transaction to let the other one succeed,
or we can also wait until the other transaction committed. The latter
is an interesting option if we are trying to commit a write and
another transaction already read the object. We can then signal the
other transaction to commit as soon as possible and wait. After
waiting, there is now no conflict between our write and the already
committed read anymore.

@@

We evaluate our system in a Python interpreter called
PyPy~\cite{pypy} (version 2.3). PyPy is an implementation of an
interpreter for the Python language. It has a special focus on speed
and provides a just-in-time (JIT) compiler to speed up applications
running on top of it. For comparison, we compare between normal PyPy
using a GIL and a PyPy with STM, as well as to CPython.
\begin{description}
\item[CPython] (version 2.7.6) is the reference implementation of the Python
  language. It is the most widely used interpreter for this language.
  The implementation uses a GIL for synchronisation in multi-threaded
  execution and it does not feature a JIT compiler.
\end{description}

@@

For each point in the plots, we took 5 measurements and report the
average and the standard deviation as error bars. For the JIT
benchmarks, we first let the JIT warm up by doing a few
additional iterations to get a single measurement.

@@

To asses how well the STM system scales on its own (without any real
workload), we execute the following loop on 1 to 4 threads on our
Python interpreter with STM:
\begin{lstlisting}
def workload():
    i = 20000000
    while i:
        i -= 1
\end{lstlisting}

For the results in Figure~\ref{fig:scaling}, we averaged over 5 runs
and normalised the average runtimes to the time it took on a single
thread. From this we see that there is additional overhead introduced
by each thread ($12.3\%$ for all 4 threads together). While this
is not ideal, we think that $12.3\%$ are acceptable on four
threads. In terms of throughput, 4 threads have $3.54\times$
more iterations per second than a single thread.

@@

Looking at the average overhead from switching from GIL to STM, we see
that it is $\approx 43.3\%$. The maximum in richards is
$71\%$. In all benchmarks \emph{pypy-stm-nojit} beats
\emph{pypy-nojit} already on two threads despite of this overhead.
However, it rarely beats CPython's single-thread performance.  For
programs that need concurrency in CPython and that use threads to
achieve this, it also makes sense to look at the overhead induced by
the GIL on multiple threads. From this perspective, the STM
implementation beats CPython's performance in all but two benchmarks.

@@

We see that generally, the group of embarrassingly parallel benchmarks scales
best. The other three scale barely or not at all with the number of
threads. The reason for this is likely again the conflicts in the
latter group. Since the JIT accelerates all code but not the STM
overhead, we do more work per transaction. This increases the
likelihood of conflicts between them and therefore limits scalability
even more than in the no-JIT benchmarks.

@@

Similarly to us, \cite{martin09} use memory management techniques to
provide isolation between transactions. They focus on providing strong
atomicity between transactional and non-transactional code.  We,
however, use inevitable transactions to provide strong atomicity
for non-reversible operations and otherwise do not have to deal
with non-transactional code.  Additionally, while they use page fault
handlers to detect conflicts, we use traditional write barriers. There
are only some superficial similarities between our approach and
theirs.

@@

\acks
We would like to thank the following people for their valuable
inputs and the many fruitful discussions: Carl F. Bolz, Michael Faes,
Pascal Felber, Patrick Marlier, and Aristeidis Mastoras.

@@

%%% TODO: fix all citation to follow common format (always page numbers; no 'et al', etc)

>>>>

% Keine "Schusterjungen"
\clubpenalty = 10000
% Keine "Hurenkinder"
\widowpenalty = 10000 \displaywidowpenalty = 10000

% \newcommand{\mynote}[2]{%
%   \textcolor{red}{%
%     \fbox{\bfseries\sffamily\scriptsize#1}%
%     {\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}%
%   }%
% }

% \newcommand\remi[1]{\mynote{Remi}{#1}}
% \newcommand\cfbolz[1]{\mynote{cfbolz}{#1}}
% \newcommand\arigo[1]{\mynote{arigo}{#1}}

% \newcommand{\comment}[1]{}

@@

% \comment{
% A Platform for Parallelism in Dynamic languages
% Parallelism for Python
% A TM Implementation for Python
% Transactional Memory as a foundation for parallel Python
% Transactional Memory as a foundation for parallelism in Python
% Parallel Python based on VM-assisted TM
% Parallel Python based on VM-assisted Transactional Memory
% A Platform for parallel Python
% A TM Platform for parallel Python
% A Transactional Memory Platform for parallel Python
% A Transactional Memory system for parallel Python

% Virtual Memory Assisted Transactional Memory for Dynamic Languages}

@@

This paper makes the following contributions:
\vspace{-\topsep}               % XXX:HACK:TODO
\begin{itemize}
\item We introduce a new STM system that performs well for dynamic
  language interpreters.
\item We integrate the system closely with a garbage collector
  (GC) to lower the overhead of STM.
\item This new STM system is used to replace the GIL in one
  implementation of Python and is then evaluated.
\end{itemize}

The STM described here is efficient and has low overhead and is
therefore beneficial also for environments with a small number of
threads or cores.

@@

@@

\begin{code}[h]
\begin{lstlisting}
void stm_start_transaction()
void stm_commit_transaction()
void stm_become_inevitable()
void stm_read(object_t *SO)
void stm_write(object_t *SO)
object_t *stm_allocate(size_t size_rounded)
STM_PUSH_ROOT(object_t *SO)
STM_POP_ROOT(object_t *SO)
\end{lstlisting}
\caption{Application programming interface\label{lst:api}}
\end{code}

Our TM system is designed as a library that covers all aspects around
transactions and object memory management. It is designed for
object-oriented dynamic language VMs as a replacement for the GIL.
The library exposes the functions in Listing~\ref{lst:api} as its
interface for use by a dynamic language interpreter.

@@

The next step involves using \lstinline!remap_file_pages()!, a Linux
system call\footnote{remap\_file\_pages() can also be done with more mmap()
calls in a roughly portable way, but the former is more efficient on Linux.},
to establish the \emph{fully-shared configuration}.  All pages
of segments $>0$ map to the pages of the sharing-segment.
However, the layout of a segment is not uniform and we actually
privatise a few areas again right away. These areas are illustrated in
Figure~\ref{fig:Segment-Layout} and explained here:

\begin{description}
\item [{NULL~page:}] This page is unmapped and will produce a
  segmentation violation when accessed. We use this to detect
  erroneous dereferencing of \lstinline!NULL! references.  All
  $\%gs{::}SO$ translated to linear addresses will point to NULL pages
  if $SO$ is set to \lstinline!NULL!.
\item [{Segment-local~data:}] Some area private to the segment that
  contains segment-local information for bookkeeping.
\item [{Read~markers:}] These are private pages that store information about
  which objects were read in the current transaction running in this
  segment.
\item [{Nursery:}] This private area contains all the freshly allocated
  objects (\emph{young objects}) of the current transaction. The GC
  uses bump-pointer allocation in this area to allocate objects in the
  first generation.
\item [{Old~object~space:}] These pages are the ones that are really
  shared between segments. They mostly contain old objects but also
  some young ones that were too big to be allocated in the nursery.
\end{description}

Note, these areas are at offsets inside the segments known at
compile time. This makes some checks very efficient, e.g.\ checking
if an object resides in the nursery only requires comparing its $SO$
to the boundaries of the nursery.

@@

At the start of a transaction, the thread needs to acquire a
segment. It may have to wait until another thread finishes its
transaction and releases its segment. Fairness is not guaranteed yet, as
we simply assume a fair scheduling policy in the operating system when
waiting on a lock.

@@

To improve this situation even more, we introduce the concept of
\emph{overflow objects}. If a minor collection needs to occur during a
transaction, we empty the nursery and mark each surviving object in
the old object space with an \lstinline!overflow_number!  globally
unique to the current transaction. That way we can still detect in a
medium-fast path inside the write barrier that the object still belongs to the
current transaction. Other and later transactions will treat them
like regular \emph{old} objects.

@@

This area can be seen as a continuous, segment-local array of bytes
that is indexed with an object's reference ($SO$) divided by 16. So
that each object has its own read marker, all objects have a size of
at least 16 bytes. Otherwise there could be false conflicts when
reading from two adjacent objects that share a single marker.
Instead of just setting the byte to \lstinline!true!  if the
corresponding object was read, we set it to a \lstinline!read_version!
belonging to the transaction, which will be incremented on each
commit.  Thereby, we can avoid resetting the bytes to
\lstinline!false!  on commit and only need to do this every 255
transactions. The whole code for the barrier shown in
Listing~\ref{lst:rb} is easily optimisable for compilers as well as
perfectly predictable for CPUs\footnote{Additional benefit: The read
barrier is not constrained to execute before the actual read -- both
the compiler and the CPU are free to reorder or group them.}.

\begin{code}[h]
\begin{lstlisting}
void stm_read(SO):
    *(SO >> 4) = read_version
\end{lstlisting}
\caption{The complete read barrier\label{lst:rb}}
\end{code}

@@

The job of the write barrier is twofold: it serves as a write barrier
for the garbage collector (GC) and the STM system. These two jobs are
depicted in Figure~\ref{fig:obj_states}. The GC needs to keep track of
objects that may contain references to young, movable objects. These
objects need to be traced during minor collections to update
references to moved objects. Young objects automatically get traced if
the GC ascertains that they survive the collection by being reachable
from traced objects. The STM system needs to record objects in the
write set, acquire write locks, and privatise their pages
(copy-on-write). It only needs to do this for old objects from
previous transactions, as the others are not known to other
transactions anyway.

The \textbf{fast path} of the write barrier is very simple. We only
need to check for the flag \lstinline!WRITE_BARRIER! in the object's
header through its $SO$ and call the slow path if it is set. This flag
is set either if the object is old and comes from an earlier
transaction, or if there was a minor collection which will add the
flag again on all objects, including new overflow objects. The flag
is never set on young, freshly allocated objects. This fast path
covers all cases for the GC and for the STM system (see
Listing~\ref{lst:w_fast}).

\begin{code}[h]
\begin{lstlisting}
void stm_write(SO):
	if SO->flags & WRITE_BARRIER:
		write_slowpath(SO)
\end{lstlisting}
\caption{Write barrier fast path\label{lst:w_fast}}
\end{code}

@@

The \textbf{slow path} is shown in Listing~\ref{lst:w_slow}.
First comes the \emph{GC part}: In any case, the object will be added
to the list of objects that need tracing in the next minor collection
(\lstinline!to_trace!). Any reference we write to this object will
be kept alive and up-to-date during minor collections.

@@

For the \emph{STM part}, we first perform a read barrier on the
object. We then try to acquire its write lock. \lstinline!write_locks!
is a simple, \emph{global} array of bytes that is indexed with the
$SO$ of the object divided by 16. Note, ``global'' here really means
it is a single array with data for all segments, there is no address
translation going on to access its elements contrary to e.g.\ the
read markers array.  If we already own the lock, we are done.
If someone else owns the lock, we perform  write-write contention
management that aborts either us or the current owner of the
object.  If we succeed in acquiring the lock using an atomic
\lstinline!cmp_and_swap!, we need to add the object to the write set
(a simple list called \lstinline!modified_old_objects!)  and privatise
all pages belonging to it (copy-on-write).

@@

Note that we have three kinds of objects: \emph{young, overflow} and
\emph{old} objects. Young and overflow objects were created in the
same transaction; old objects always come from previous transactions.
The close integration between STM and the GC allows us to use this
information effectively: young objects do not trigger the slow path at
all, overflow objects only need to be handled by the GC part of the
write barrier once between collections, and only the old objects need
to be handled by the STM part \emph{once per transaction}. This keeps the
overhead of the write barrier very low.

@@

A requirement for performing a commit is to synchronise all threads so
that we can safely update objects in other segments. To make this
synchronisation fast and cheap, we do not want to insert an additional
check regularly to see if synchronisation is requested. We
use a trick relying on the fact that dynamic languages are usually
very high-level and thus allocate a lot of objects very regularly.
This is done through the function \lstinline!stm_allocate!  shown
in Listing~\ref{lst:alloc_func}.

\begin{code}[h]
\begin{lstlisting}
object_t *stm_allocate(ssize_t size_rounded):
    result = nursery_current
	nursery_current += size_rounded
	if nursery_current > nursery_end:
		return allocate_slowpath(size_rounded)
	return result
\end{lstlisting}
\caption{Function to allocate objects\label{lst:alloc_func}}
\end{code}

@@

On encountering conflicts, we employ contention management to solve
the problem as well as we can. The general rules are:

\begin{itemize}
\item prefer transactions that started earlier to younger transactions
  to increase the chance of long transactions succeeding
\item to support \emph{inevitable} transactions, we always prefer them
  to others since they cannot abort (similar to~\cite{blundell06})
\end{itemize}

We can either simply abort a transaction to let the other one succeed,
or we can also wait until the other transaction committed. The latter
is an interesting option if we are trying to commit a write and
another transaction already read the object. We can then signal the
other transaction to commit as soon as possible and wait. After
waiting, there is now no conflict between our write and the already
committed read anymore.

@@

We evaluate our system in a Python interpreter called
PyPy~\cite{pypy} (version 2.3). PyPy is an implementation of an
interpreter for the Python language. It has a special focus on speed
and provides a just-in-time (JIT) compiler to speed up applications
running on top of it. We compare between normal PyPy
using a GIL and a PyPy with STM, as well as to CPython.
\begin{description}
\item[CPython] (version 2.7.6) is the reference implementation of the Python
  language. It is the most widely used interpreter for this language.
  The implementation uses a GIL for synchronisation in multi-threaded
  execution and it does not feature a JIT compiler.
\end{description}

@@

For each point in the plots, we took 5 measurements and report the
arithmetic mean with the standard deviation as error bars. Average
speedup numbers are calculated using the geometric mean. For the JIT
benchmarks, we first let the JIT warm up by doing a few additional
iterations to get a single measurement.

@@

To asses how well the STM system scales on its own (without any real
workload), we execute the loop in Listing~\ref{lst:scaling_workload}
on 1 to 4 threads on the PyPy interpreter with STM.

\begin{code}[h]
\begin{lstlisting}
def workload():
    i = 20000000
    while i:
        i -= 1
\end{lstlisting}
\caption{Dummy workload\label{lst:scaling_workload}}
\end{code}

For the results in Figure~\ref{fig:scaling}, we averaged over 5 runs
and normalised the average runtimes to the time it took on a single
thread. From this we see that there is additional overhead introduced
by each thread ($12.3\%$ for all 4 threads together). Every thread
adds some overhead because during a commit, there is one more thread
which has to reach a safe point. Additionally, conflict detection
during commit needs to check in all concurrently running transactions.

While not ideal, we think that $12.3\%$ is acceptable on four
threads. In terms of throughput, 4 threads have $3.54\times$
more iterations per second than a single thread.

@@

Looking at the average overhead from switching from GIL to STM, we see
that it is $\approx 43.3\%$. The maximum in richards is $71\%$. In all
benchmarks \emph{pypy-stm-nojit} beats \emph{pypy-nojit} already on
two threads despite of this overhead.  The achieved speedup comparing
STM to the GIL is between $1.14\times$ and $1.94\times$.

Still, STM rarely beats CPython's single-thread performance. However, for
programs that need concurrency in CPython and that use threads to
achieve this, it also makes sense to look at the overhead induced by
the GIL on multiple threads. From this perspective, the STM
implementation beats CPython's performance in all but two benchmarks.

@@

We see that generally, the group of embarrassingly parallel benchmarks
scales best. (There is a notable performance stability problem in the
\emph{richards} benchmark. This is a bug we will fix in the future.)
The other three benchmarks scale barely or not at all with the number of
threads. The reason for this is likely again the conflicts in the
latter group. Since the JIT accelerates all code but not the STM
overhead, we do more work per transaction. This increases the
likelihood of conflicts between them and therefore limits scalability
even more than in the no-JIT benchmarks.

@@

Similarly to us, \cite{martin09} use memory management techniques to
provide isolation between transactions. They focus on providing strong
atomicity between transactional and non-transactional code.  We,
however, use inevitable transactions to provide strong atomicity
for non-reversible operations and otherwise do not have to deal
with non-transactional code.  Additionally, while they use page fault
handlers to detect conflicts, we use traditional write barriers. There
are only some superficial similarities between the approach presented
here and theirs.

@@

\acks
We would like to thank the following people for their valuable
inputs and the many fruitful discussions: Carl F. Bolz, Michael Faes,
Pascal Felber, Maciej Fijalkowski, Patrick Marlier, and Aristeidis Mastoras.

@@
